---
title: "Forest Growth with Sobol and ODE"
author: "Mallory Giesie"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)
library(ggplot2)
```

**Running our forest growth model for 300 years using ODE solver**

**Our Parameters:**

*-thresholdhold = 50 kgC*

*-K = 250 kg C (carrying capacity)*

*-r = 0.01 (exponential growth rate before before canopy closure)*

*-g = 2 kg/year (linear growth rate after canopy closure)*

```{r}
#sourcing in our forest size calculating model
source(here("dforestgrowth.R"))
```

```{r 300 years}
C = 10 #initial size of forest

parms_initial = c(list(K=250, r=0.1, g=2, thresholdhold = 50))

simtimes = seq(from=1, to=300)

result = ode(y=C, times=simtimes, func=dforestgrowth, parms=parms_initial)

head(result) #check first rows of result

colnames(result)=c("time","C") #adjust column names

# turn it into a data frame
result = as.data.frame(result)
```

**Graphing 300 year trajectory**

```{r}
ggplot(result, aes(x = time, y = C)) +
  geom_point() +
  ggtitle("300 year forest growth") +
  theme_minimal()
```

**Runing a sobol global sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of 300 years, varies with these parameters:**

*-pre canopy closure growth rate (r)*

*-post-canopy closure growth rate (g)*

*-canopy closure threshold & carrying capacity(K)*

```{r setting up parameter values}
# first set of samples
np=2000

K = rnorm(mean=250, sd=25, n=np)

r = rnorm(mean=0.01, sd=0.001, n=np)

g = rnorm(mean=2, sd=0.2, n=np)

threshold = rnorm(mean=50, sd=5, n=np)

X1 = cbind.data.frame(K=K, r=r, g=g, threshold = threshold)

# repeat to get our second set of samples
K = rnorm(mean=100, sd=10, n=np)

r = rnorm(mean=0.01, sd=0.001, n=np)

g = rnorm(mean=2, sd=0.2, n=np)

thresholdh = rnorm(mean=50, sd=5, n=np)

X2 = cbind.data.frame(K=K, r=r, g=g, threshold = threshold)

# fix any negative values
X1 = X1 %>% map_df(pmax, 0.0) #running each parameter through pmax(carrying capacity and growth rates)

X2 = X2 %>% map_df(pmax, 0.0)

# Creating our sobel object
sens_C = sobolSalt(model = NULL,X1, X2, nboot = 300) 

head(sens_C$X)

# lets add names
colnames(sens_C$X) = c("K","r", "g", "threshold")
```

```{r computing max pop metrics}
# function to compute metrics 
compute_metrics = function(result) {
  maxpop = max(result$C)
  return(list(maxpop = maxpop))
}

compute_metrics(result)


c_wrapper = function(K,
                     r,
                     g,
                     threshold,
                     C,
                     simtimes,
                     func) {
  parms = list(
    K = K,
    r = r,
    g = g,
    threshold = threshold
  )
  
  result = ode(
    y = C,
    times = simtimes,
    func = func,
    parms = parms
  )
  
  colnames(result) = c("time", "C")
  
  # getting metrics
  
  metrics = compute_metrics(as.data.frame(result))
  
  return(metrics)
}

# now use pmap as we did before
allresults = as.data.frame(sens_C$X) %>% pmap(c_wrapper,
                                              C = C,
                                              simtimes = simtimes,
                                              func = dforestgrowth)


# extract out results from pmap into a data frame
allres = allresults %>% map_dfr(`[`,c("maxpop"))


# create boxplots
tmp = allres %>% pivot_longer(cols=everything(),names_to="metric", values_to="value")
```

**Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T)** 

```{r box plot}
ggplot(tmp, aes(metric, value)) +
  geom_boxplot(fill = "lightblue") +
  theme_minimal() +
  labs(title = "Sensitivity Analysis of Maximum Forest Size",
       y = "Max Forest Size (kgC)")
```

```{r sobol indices}
sens_C_maxpop = sensitivity::tell(sens_C,allres$maxpop)

sens_C_maxpop$S

sens_C_maxpop$T
```

